---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.align = 'center', message = FALSE, 
                      warning = FALSE, fig.width = 8, fig.height = 8)
```

`r Sys.Date()`

# Coverage Simulation

Ran 

- 100 simulations with 
- 25 subjects, each going through
- 10, 25, 50, 75, and 100 trials with
- 1000 bootstraps each

Coverage is proportion of simulations in which the true curve was between the 5th and 95th quantiles of the bootstrapped distribution

# Parameters

```{r}

library(data.table)
library(bdots)
library(eyetrackSim)
library(mvtnorm)

results <- readRDS("../data/coverage_results.rds")
res_bdots <- readRDS("../data/bdots_coverage_results.rds")

pp <- lapply(results, `[[`, 1)
ppb <- lapply(res_bdots, `[[`, 1)
cc <- lapply(results, `[[`, 2)
ccb <- lapply(res_bdots, `[[`, 2)

## Parameters first, let's say 90 percent coverage
getParCoverage <- function(pp) {
  pp <- lapply(pp, function(x) {
    parCoverage <- apply(x, 2, function(y) sum(y > 0.05 & y < 0.95) / length(y))
  })
  
  mini <- lapply(pp, `[[`, 1) |> unlist()
  peak <- lapply(pp, `[[`, 2) |> unlist()
  slope <- lapply(pp, `[[`, 3) |> unlist()
  cross <- lapply(pp, `[[`, 4) |> unlist()
  
  return(list(mini = mini, peak = peak, slope = slope, cross = cross))
}

pp <- getParCoverage(pp)
ppb <- getParCoverage(ppb)

par(mfrow = c(2,2))
plot(pp$mini, ylim = c(0, 1.1), main = "mini", type = 'b', xaxt = 'n', 
     xlab = "# Trials", ylab = "Coverage", pch = 1)
abline(h = 0.9, col = 'red', lty = 2)
axis(1, at = 1:5, labels = c(10, 25, 50, 75, 100))
lines(ppb$mini, type = 'b', pch = 2, col = 'blue')
legend(1, 0.2, legend = c("Bootstrap", "bdots"), pch = c(1, 2),
       col = c("black", "blue"))

plot(pp$peak, ylim = c(0, 1.1), main = "peak", type = 'b', xaxt = 'n', 
     xlab = "# Trials", ylab = "Coverage")
abline(h = 0.9, col = 'red', lty = 2, ylab = "Coverage")
axis(1, at = 1:5, labels = c(10, 25, 50, 75, 100))
lines(ppb$peak, type = 'b', pch = 2, col = 'blue')


plot(pp$slope, ylim = c(0, 1.1), main = "slope", type = 'b', xaxt = 'n', 
     xlab = "# Trials", ylab = "Coverage")
abline(h = 0.9, col = 'red', lty = 2, ylab = "Coverage")
axis(1, at = 1:5, labels = c(10, 25, 50, 75, 100))
lines(ppb$slope, type = 'b', pch = 2, col = 'blue')

plot(pp$cross, ylim = c(0, 1.1), main = "cross", type = 'b', xaxt = 'n', 
     xlab = "# Trials", ylab = "Coverage")
abline(h = 0.9, col = 'red', lty = 2, ylab = "Coverage")
axis(1, at = 1:5, labels = c(10, 25, 50, 75, 100))
lines(ppb$cross, type = 'b', pch = 2, col = 'blue')
```

# Average pointwise coverage by trial

```{r, fig.width=6, fig.height=6}
## This takes average across time points, then average across trials, then average total
cc_cover <- lapply(cc, function(x) {
  ## Here we just do full band for now
  mm <- apply(x, 1, function(y) sum(y > 0.05 & y < 0.95) / length(y))
  mean(mm)
}) |> unlist()

## Do this also for bdots, which only has binary indicators of coverage rather than quantile
cc_coverb <- lapply(ccb, function(x) {
  ## Here we just do full band for now
  mm <- apply(x, 1, function(y) mean(y))
  mean(mm)
}) |> unlist()


#cc_cover <- Reduce(`+`, cc_cover) / length(cc_cover)
par(mfrow = c(1,1))
plot(cc_cover, ylim = c(0.0, 1.1), main = "Average pointwise coverage of curve", type = 'b',
     xaxt = 'n', xlab = "# Trials", ylab = "Coverage")
abline(h = 0.9, col = 'red', lty = 2)
axis(1, at = 1:5, labels = c(10, 25, 50, 75, 100))
lines(cc_coverb, type = 'b', pch = 2, col = 'blue')
legend("bottomleft", legend = c("bootstrap", "bdots"), pch = c(1, 2), col = c("black", "blue"))
```

# Average pointwise coverage by time

```{r, fig.width=12}

## Here we just do full band for now
cc_time <- lapply(cc, function(x) {
  mm <- apply(x, 1, function(y) sum(y > 0.05 & y < 0.95) / length(y))
}) 
cc_timeb <- lapply(ccb, function(x) {
  mm <- apply(x, 1, function(y) mean(y))
}) 

par(mfrow = c(2, 3))
mtrials <- c(10, 25, 50, 75, 100)
for (i in seq_along(cc_time)) {
  x <- cc_time[[i]]
  y <- cc_timeb[[i]]
  plot(x, ylim = c(0.0, 1.05), xlab = "Time", xaxt = 'n', type = 'l',
       ylab = 'coverage', main = paste(mtrials[i], "Trials"), lwd = 1.2)
  abline(h = 0.9, col = 'red', lty = 2)
  axis(1, at = seq(0, 501, by = 12)[-2], labels = seq(0, 2000, by = 50))
  lines(y, lwd = 1.2, col = 'blue')
  legend("bottomleft", legend = c("bootstrap", "bdots"), col = c("black", "blue"), lty = 1)
}

```


# Quantile needed for 90% band coverage

We don't have this for bdots as bdots does not naturally produce quantiles. Instead, here we set $\alpha = 0.1$ and considered pointwise coverage to be whether or not the generating curve fell within the $\mu_t \pm t_{\alpha^*}\hat{\sigma}_t$ band. Since the bootstraps do appear to generally be normally distributed at each point, this range is actually more generous than simply having determined the collection of bootstrapped curves, measuring the quantiles from there. 

```{r}

cc_band <- lapply(cc, function(x) {
  ## Go through each trial and determine what quantile
  # needed to cover entire band
  mm <- apply(x, 2, function(y) max(max(y), 1 - min(y)))
  quantile(mm, 0.9)
})

par(mfrow = c(1,1))
plot(unlist(cc_band), ylim = c(0.8, 1.0), main = "Required quantile for 90% coverage of entire band", type = 'b',
     xaxt = 'n', xlab = "# Trials", ylab = "Coverage")
abline(h = 0.9, col = 'red', lty = 2)
axis(1, at = 1:5, labels = c(10, 25, 50, 75, 100))
text(1:5, unlist(cc_band) + 0.005, unlist(cc_band), cex = 0.8)
```
