---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.align = 'center', fig.height = 4, fig.width = 6)
library(ggplot2)
library(bdots)
library(gridExtra)
```

# Irl data

6 way comparison of the 4 groups in bob's sli paper. Only two group comparisons yielded any differences

## set 1

```{r, include = FALSE}
load("~/dissertation/analysis/irl_comparison/irl_comparison.rds")

p1 <- plot(boot_l2, plotDiffs = FALSE)$bootPlot 
p2 <- plot(boot_s2, plotDiffs = FALSE)$bootPlot 
p1 <- p1 + ggtitle("Bootstrapped Fits -- Proportion of Fixations")
p2 <- p2 + ggtitle("Bootstrapped Fits -- Look Onset")
```

```{r, fig.width = 10}
grid.arrange(p1, p2, nrow = 1)
```

## set 2

```{r, include=FALSE}
p3 <- plot(boot_l4, plotDiffs = FALSE)$bootPlot
p4 <- plot(boot_s4, plotDiffs = FALSE)$bootPlot

p3 <- p3 + ggtitle("Bootstrapped Fits -- Proportion of Fixations")
p4 <- p4 + ggtitle("Bootstrapped Fits -- Look Onset")
```

```{r, fig.width = 10}
grid.arrange(p3, p4, nrow = 1)
```

# power with onset vs proportion

In this first batch (with 100 simulations), changed logistic peak from 0.8 to 0.885 and changed double gauss mean from 630 to 700

```{r, echo = FALSE, fig.align='center'}
powerHist <- function(y, tit) {
  y <- readRDS(y)
  ons <- y[1, ]
  fix <- y[2, ]

  timesig <- function(mm) {
    time <- seq(0, 2000, 4)
    vec <- vector("numeric", length = length(time))

    if (is.null(dim(mm))) {
      vec <- as.logical(vec)
      return(vec)
    }

    sm <- split(mm, row(mm))
    bv <- lapply(sm, function(m) {
      sigt <- do.call(seq, as.list(c(m, 4)))
    })
    bv <- Reduce(union, bv)
    rr <- time %in% bv # logical vec
    return(rr)
  }

  TIME <- seq(0, 2000, 4)

  ons_mat <- sapply(ons, timesig)
  ons_mat <- apply(ons_mat, 2, function(z) TIME[z])
  ons_mat <- Reduce("c", ons_mat)

  fix_mat <- sapply(fix, timesig)
  fix_mat <- apply(fix_mat, 2, function(z) TIME[z])
  fix_mat <- Reduce("c", fix_mat)

  s1 <- hist(ons_mat, breaks = seq(0, 2000, 40), plot = FALSE)$counts
  s2 <- hist(fix_mat, breaks = seq(0, 2000, 40), plot = FALSE)$counts

  yy <- max(s1, s2)
  hist(ons_mat, breaks = seq(0, 2000, 40), ylim = c(0, yy), main = tit,
       xlab = "Time", col = alpha("steelblue", alpha = 0.5))
  hist(fix_mat, breaks = seq(0, 2000, 40), ylim = c(0, yy), main = tit,
       xlab = "Time", col = alpha("tomato", alpha = 0.5), add = TRUE)
}

ff <- list.files("~/dissertation/writing/saccade/scripts/sim_bdots_power_methods/rds_files", full.names = TRUE, pattern = "rds")




powerHist(ff[1], "no delay lg")
powerHist(ff[2], "norm delay lg")
powerHist(ff[3], "weib delay lg")

powerHist(ff[4], "no delay dg")
powerHist(ff[5], "norm delay dg")
powerHist(ff[6], "weib delay dg")

```

In this second batch (N = 250), I changed logistic crossover 765 to 700 and left dg same as before (not all simulations of this are finished)

```{r, echo = FALSE, fig.align='center'}
ff <- list.files("~/dissertation/writing/saccade/scripts/sim_bdots_power_methods/alt_rds_files", full.names = TRUE, pattern = "rds")

powerHist(ff[1], "norm delay lg")
powerHist(ff[2], "weib delay lg")
powerHist(ff[3], "norm delay dg")
```


# Logistic shift

try to determine power between these two methods in case of 100ms and  200ms shift in crossover with differing levels of variability (need to do that 4-way analysis but haven't)

```{r, include=FALSE}

ff <- list.files("~/dissertation/writing/methodology/scripts/logistic_power/rds_files/", full.names = TRUE, pattern = "rds")

## Given signifiance mat, need to return length 401 bool vector
timetiePower <- function(mm) {
  time <- seq(0, 1600, 4)
  vec <- vector("numeric", length = length(time))

  if (is.null(dim(mm))) {
    vec <- as.logical(vec)
    return(vec)
  }

  # each of these into a list
  sm <- split(mm, row(mm))
  bv <- lapply(sm, function(m) {
    #sigt <- do.call(seq, as.list(c(m, by = 0.005)))
    time[time >= m[1] & time <= m[2]] # for rounding
  })
  bv <- Reduce(union, bv)
  rr <- time %in% bv #round(time,3) %in% round(bv, 3)
  return(rr)
}




getDiffSlicesgg <- function(ff, ww, leg = TRUE) {
  rr <- readRDS(ff)

  tt <- attributes(rr)[[1]] #|> unlist()

  vs <- ifelse(tt$pm == 1, "\nLow Var", "\nHigh Var")

  tit <- paste0("Paired: ", tt$paired, vs, "\nShift: ", tt$shift)
  sm <- lapply(rr, `[[`, 1)
  mm <- lapply(rr, `[[`, 2)
  pm <- lapply(rr, `[[`, 3)

  smm <- sapply(sm, timetiePower) |> rowSums()
  mmm <- sapply(mm, timetiePower) |> rowSums()
  pmm <- sapply(pm, timetiePower) |> rowSums()
  time <- seq(0, 1600, 4)

  dat <- data.table(Method = rep(c("Hom. Boot",
                                   "Het. Boot",
                                   "Permutation"), each = 401),
                    Power = c(smm, mmm, pmm),
                    Time = rep(time, 3))

  ll <- length(rr)

  pp <- ggplot(dat, aes(Time, Power, color = Method)) + theme_bw() + ggtitle(tit) +
    geom_line(linewidth=1) + #geom_abline(slope = 0, intercept = 5, color = 'red', linetype = "dotted")
    scale_color_manual(values = c("#00BFC4", "#7CAE00", "#C77CFF")) + theme(legend.position = "bottom")

  if (!leg) pp <- pp + theme(legend.position = "none")
  return(pp)
}

## Shift of 100
p1 <- getDiffSlicesgg(ff[1], 1)
p2 <- getDiffSlicesgg(ff[2], 2)
p3 <- getDiffSlicesgg(ff[3], 3)
p4 <- getDiffSlicesgg(ff[4], 4)

## Shift of 200
p5 <- getDiffSlicesgg(ff[5], 5)
p6 <- getDiffSlicesgg(ff[6], 6)
p7 <- getDiffSlicesgg(ff[7], 7)
p8 <- getDiffSlicesgg(ff[8], 8)
```

```{r, fig.width=10, fig.height=10}
grid.arrange(p1, p2, p3, p4, nrow = 2)

grid.arrange(p5, p6, p7, p8, nrow = 2)
```


# Tables

Coming back to FWER now that I have the paired permutation "fixed". Not sure exactly yet but will explore on my own time

```{r}
library(xtable)

getFWER <- function(y) {
  sigs <- readRDS(y)

  sm <- sapply(sigs, function(x) !is.null(x[[1]])) |> mean()
  mm <- sapply(sigs, function(x) !is.null(x[[2]])) |> mean()
  pm <- sapply(sigs, function(x) !is.null(x[[3]])) |> mean()

  smt <- sapply(sigs, function(x) timetie(x[[1]])) |> rowMeans()
  mmt <- sapply(sigs, function(x) timetie(x[[2]])) |> rowMeans()
  pmt <- sapply(sigs, function(x) timetie(x[[3]])) |> rowMeans()

  fwer <- data.table(sm = sm, mm = mm, pm = pm)
  tsmat <- matrix(c(smt, mmt, pmt), byrow = TRUE, nrow = 3,
                  dimnames = list(c("sm", "mm", "pm"), NULL))
  return(list(fwer = fwer, timeSliceMat = tsmat))
}

## This takes the list returned by getfwer
timeSliceFwer <- function(y, f = median) {
  nn <- length(y)
  y <- lapply(y, `[[`, 2)
  y <- lapply(y, function(z) {
    apply(z, 1, f)
  })
  y <- Reduce(rbind, y)
  #y <- sapply(y, rowMeans)
  #rownames(y) <- paste0("sim", 1:8)
  y
}

## Given signifiance mat, need to return length 401 bool vector
timetie <- function(mm) {
  time <- seq(0, 1600, 4)
  vec <- vector("numeric", length = length(time))

  if (is.null(dim(mm))) {
    vec <- as.logical(vec)
    return(vec)
  }

  # each of these into a list
  sm <- split(mm, row(mm))
  bv <- lapply(sm, function(m) {
    sigt <- do.call(seq, as.list(c(m, 4)))
  })
  bv <- Reduce(union, bv)
  rr <- time %in% bv
  return(rr)
}


ff <- list.files("~/dissertation/writing/methodology/scripts/argon/feb_rds_files", full.names = TRUE, pattern = "rds")
#ff <- ff[c(1, 5:12, 2:4)]
ff <- ff[c(1, 9:16, 2:8)]

# don't have manymeans false paired true
ff <- ff[c(1:8, 13:16)]


sds <- expand.grid(bdotscor = c(TRUE, FALSE),
                   ar1 = c(TRUE, FALSE),
                   manymeans = c(FALSE, TRUE),
                   paired = c(FALSE, TRUE))
sds <- as.data.table(sds)

if (length(ff) != 16) {
  sds <- sds[!(manymeans == FALSE & paired == TRUE), ]
}




gg <- lapply(ff, getFWER)
fwer <- rbindlist(lapply(gg, `[[`, 1))
tfwer <- timeSliceFwer(gg, median)

fwer <- as.data.table(cbind(sds, fwer))[order(rev(manymeans), ar1, decreasing = TRUE), ]
tfwer <- as.data.table(cbind(sds, tfwer))[order(rev(manymeans), ar1, decreasing = TRUE)]

fwer <- fwer[order(manymeans), ]
tfwer <- tfwer[order(manymeans), ]

fwers <- split(fwer, by = "paired")
tfwers <- split(tfwer, by = "paired")


fwers[[2]][, c(1,3:7)] #|> xtable() |> print(include.rownames = FALSE)
# Unpaired
fwers[[1]][, c(1,3:7)] #|> xtable() |> print(include.rownames = FALSE)
```
