---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center', 
                      fig.height = 8, fig.width = 8, 
                      warning = FALSE, message = FALSE)
```

`r Sys.Date()`


Basically just a sanity check to verify that I can indeed recover the true curve if I were to observe it exactly

Maybe try this using few time points

## logistic

```{r, message=FALSE}
library(eyetrackSim)
library(bdots)

ci <- as.data.table(ci)
ci <- ci[LookType == "Target" & Subject %in% 2:6, ]

fit <- bdotsFit(data = ci, 
                subject = "Subject", 
                time = "Time", 
                y = "Fixations", 
                group = "protocol", 
                curveType = logistic())

plot(fit)

## Get realistic parameter values
mm <- coef(fit)

## Create new fixations based on pars
newdat <- Map(function(x, y) {
  x[, Fixations := eyetrackSim:::logistic_f(y, Time)]
}, split(ci, by = "Subject"), split(mm, row(mm)))
newdat <- rbindlist(newdat)

fit <- bdotsFit(data = newdat, 
                subject = "Subject", 
                time = "Time", 
                y = "Fixations", 
                group = "protocol", 
                curveType = logistic())

plot(fit)

coef(fit)
mm
```

## dg

```{r}
dat <- as.data.table(cohort_unrelated)
dat <- dat[Subject %in% 1:4 & Group == "50" & LookType == "Cohort", ]

fit <- bdotsFit(data = dat, 
                subject = "Subject", 
                time = "Time", 
                y = "Fixations", 
                group = "Group", 
                curveType = doubleGauss(concave = TRUE))
plot(fit)

## Get realistic parameter values
mm <- coef(fit)

## Create new fixations based on pars
newdat <- Map(function(x, y) {
  x[, Fixations := eyetrackSim:::doubleGauss_f(y, Time)]
}, split(dat, by = "Subject"), split(mm, row(mm)))
newdat <- rbindlist(newdat)

fit <- bdotsFit(data = newdat, 
                subject = "Subject", 
                time = "Time", 
                y = "Fixations", 
                group = "Group", 
                doubleGauss(concave = TRUE))
plot(fit)
coef(fit)
mm
```

# Ok, but if we do different time densities

### 250 time points

```{r}
ci <- as.data.table(ci)
ci <- ci[LookType == "Target" & Subject %in% 2:6, ]

fit <- bdotsFit(data = ci, 
                subject = "Subject", 
                time = "Time", 
                y = "Fixations", 
                group = "protocol", 
                curveType = logistic())

## Get realistic parameter values
mm <- coef(fit)

## Create new fixations based on pars
newdat <- Map(function(x, y) {
  x[, Fixations := eyetrackSim:::logistic_f(y, Time)]
}, split(ci, by = "Subject"), split(mm, row(mm)))
newdat <- rbindlist(newdat)

## But now lets spruce up the time sample (length 501)
wts <- exp(-(1:501/501))
TIME <- unique(ci$Time)
earlyT <- sort(sample(TIME, 250, prob = wts))
lateT <- sort(sample(TIME, 250, prob = rev(wts)))

tt <- split(newdat, by = "Subject")
tt[[1]] <- tt[[1]][Time %in% earlyT]
tt[[2]] <- tt[[2]][Time %in% earlyT]
tt[[3]] <- tt[[3]][Time %in% lateT]
tt[[4]] <- tt[[4]][Time %in% lateT]
newdat <- rbindlist(tt)

fit <- bdotsFit(data = newdat, 
                subject = "Subject", 
                time = "Time", 
                y = "Fixations", 
                group = "protocol", 
                curveType = logistic())

## not cool with weird times
# plot(fit)
fv <- lapply(split(fit, by = "Subject"), function(x) {
  data.table(sub = x$Subject, fv = fitted.values(x$fit[[1]]))
})
fv <- rbindlist(fv)
fv[, Time := c(earlyT, earlyT, lateT, lateT)]
fv[, Fixations := newdat$Fixations]
```

Subjects 2 and 3 are early, 4 and 6 are late

```{r}
ggplot(fv, aes(x = Time, y = fv, color = sub)) +
  geom_line() + 
  geom_line(aes(x = Time, y = Fixations, color = sub), linetype = "dashed", size = 1) + 
  facet_wrap(~sub)

coef(fit)
mm
```


### 50 time points

```{r}
## Create new fixations based on pars
newdat <- Map(function(x, y) {
  x[, Fixations := eyetrackSim:::logistic_f(y, Time)]
}, split(ci, by = "Subject"), split(mm, row(mm)))
newdat <- rbindlist(newdat)

## But now lets spruce up the time sample (length 501)
wts <- exp(-(1:501/501))
TIME <- unique(ci$Time)
earlyT <- sort(sample(TIME, 50, prob = wts))
lateT <- sort(sample(TIME, 50, prob = rev(wts)))

tt <- split(newdat, by = "Subject")
tt[[1]] <- tt[[1]][Time %in% earlyT]
tt[[2]] <- tt[[2]][Time %in% earlyT]
tt[[3]] <- tt[[3]][Time %in% lateT]
tt[[4]] <- tt[[4]][Time %in% lateT]
newdat <- rbindlist(tt)

fit <- bdotsFit(data = newdat, 
                subject = "Subject", 
                time = "Time", 
                y = "Fixations", 
                group = "protocol", 
                curveType = logistic())

## not cool with weird times
# plot(fit)
fv <- lapply(split(fit, by = "Subject"), function(x) {
  data.table(sub = x$Subject, fv = fitted.values(x$fit[[1]]))
})
fv <- rbindlist(fv)
fv[, Time := c(earlyT, earlyT, lateT, lateT)]
fv[, Fixations := newdat$Fixations]
```

Subjects 2 and 3 are early, 4 and 6 are late

```{r}
ggplot(fv, aes(x = Time, y = fv, color = sub)) +
  geom_line() + 
  geom_line(aes(x = Time, y = Fixations, color = sub), linetype = "dashed", size = 1) + 
  facet_wrap(~sub)

coef(fit)
mm
```


### ok, 50 time points, mostly in the middle

```{r}
## Create new fixations based on pars
newdat <- Map(function(x, y) {
  x[, Fixations := eyetrackSim:::logistic_f(y, Time)]
}, split(ci, by = "Subject"), split(mm, row(mm)))
newdat <- rbindlist(newdat)

## But now lets spruce up the time sample (length 501)
#wts <- exp(-(1:501/501))
wts <- seq(from = -2, to = 2, length.out = length(TIME))
wts <- dnorm(wts)
earlyT <- sort(sample(TIME, 50, prob = wts))
lateT <- sort(sample(TIME, 50, prob = rev(wts)))

tt <- split(newdat, by = "Subject")
tt[[1]] <- tt[[1]][Time %in% earlyT]
tt[[2]] <- tt[[2]][Time %in% earlyT]
tt[[3]] <- tt[[3]][Time %in% lateT]
tt[[4]] <- tt[[4]][Time %in% lateT]
newdat <- rbindlist(tt)

fit <- bdotsFit(data = newdat, 
                subject = "Subject", 
                time = "Time", 
                y = "Fixations", 
                group = "protocol", 
                curveType = logistic())

## not cool with weird times
# plot(fit)
fv <- lapply(split(fit, by = "Subject"), function(x) {
  data.table(sub = x$Subject, fv = fitted.values(x$fit[[1]]))
})
fv <- rbindlist(fv)
fv[, Time := c(earlyT, earlyT, lateT, lateT)]
fv[, Fixations := newdat$Fixations]
```

Subjects 2 and 3 are early, 4 and 6 are late

```{r}
ggplot(fv, aes(x = Time, y = fv, color = sub)) +
  geom_line() + 
  geom_line(aes(x = Time, y = Fixations, color = sub), linetype = "dashed", size = 1) + 
  facet_wrap(~sub)

coef(fit)
mm
```


# Sample based on curve

Here, we will generate an exact curve, but rather than punching that into bdots, we will use it as a probabilitic function to sample new points some number of times and see what shakes (this accounts for fact that the proprotion at any time may asymptotically approach true value $f_{\theta}(t)$, but any finite collection will necessarily have some variability).

## n = 50, all time points

```{r}

## Make mini 0
mm[, "mini"] <- 0

newdat <- Map(function(x, y) {
  x[, Fixations := eyetrackSim:::logistic_f(y, Time)]
}, split(ci, by = "Subject"), split(mm, row(mm)))
# newdat <- rbindlist(newdat)

## Create new sample of observed values
newdat <- lapply(newdat, function(x, n = 50) {
  x[, probFixation := mean(rbinom(n, 1, Fixations)), by = .(Subject, Time)]
})
newdat <- rbindlist(newdat)

makenewdat <- function(x, y, nn = 50) {
  newdat <- Map(function(x, y) {
    x[, Fixations := eyetrackSim:::logistic_f(y, Time)]
  }, split(ci, by = "Subject"), split(mm, row(mm)))
  newdat <- lapply(newdat, function(x, n = nn) {
    x[, probFixation := mean(rbinom(n, 1, Fixations)), by = .(Subject, Time)]
  })
  newdat <- rbindlist(newdat)
}

newdat <- makenewdat(ci, mm, 50)

fit <- bdotsFit(data = newdat, 
                subject = "Subject", 
                time = "Time", 
                y = "probFixation", 
                group = "protocol", 
                curveType = logistic())
plot(fit)

coef(fit)
mm
```


## ok, once more, but with varying time points {.tabset}

Sampling 50 of the 501 time points

### Early and Late time

Early is subjects 2,3, late is 4,6

```{r}
# newdat <- Map(function(x, y) {
#   x[, Fixations := eyetrackSim:::logistic_f(y, Time)]
# }, split(ci, by = "Subject"), split(mm, row(mm)))
# # newdat <- rbindlist(newdat)
# 
# ## Create new sample of observed values
# newdat <- lapply(newdat, function(x, n = 50) {
#   x[, probFixation := mean(rbinom(n, 1, Fixations)), by = .(Subject, Time)]
# })
# newdat <- rbindlist(newdat)

makenewdat <- function(x, y, nn = 50) {
  newdat <- Map(function(x, y) {
    x[, Fixations := eyetrackSim:::logistic_f(y, Time)]
  }, split(ci, by = "Subject"), split(mm, row(mm)))
  newdat <- lapply(newdat, function(x, n = nn) {
    x[, probFixation := mean(rbinom(n, 1, Fixations)), by = .(Subject, Time)]
  })
  newdat <- rbindlist(newdat)
}

newdat <- makenewdat(ci, mm, 50)


wts <- seq(from = -2, to = 2, length.out = length(TIME))
wts <- dnorm(wts)
TIME <- unique(ci$Time)
earlyT <- sort(sample(TIME, 50, prob = wts))
lateT <- sort(sample(TIME, 50, prob = rev(wts)))

tt <- split(newdat, by = "Subject")
tt[[1]] <- tt[[1]][Time %in% earlyT]
tt[[2]] <- tt[[2]][Time %in% earlyT]
tt[[3]] <- tt[[3]][Time %in% lateT]
tt[[4]] <- tt[[4]][Time %in% lateT]
newdat <- rbindlist(tt)

fit <- bdotsFit(data = newdat, 
                subject = "Subject", 
                time = "Time", 
                y = "probFixation", 
                group = "protocol", 
                curveType = logistic())
# plot(fit)

fv <- lapply(split(fit, by = "Subject"), function(x) {
  data.table(sub = x$Subject, fv = fitted.values(x$fit[[1]]))
})
fv <- rbindlist(fv)
fv[, Time := c(earlyT, earlyT, lateT, lateT)]
fv[, probFixation := newdat$probFixation]
ggplot(fv, aes(x = Time, y = fv, color = sub)) +
  geom_line() + 
  geom_line(aes(x = Time, y = probFixation, color = sub), linetype = "dashed", size = 1) + 
  facet_wrap(~sub)


coef(fit)
mm
```

### Mid time

```{r}
# newdat <- Map(function(x, y) {
#   x[, Fixations := eyetrackSim:::logistic_f(y, Time)]
# }, split(ci, by = "Subject"), split(mm, row(mm)))
# # newdat <- rbindlist(newdat)
# 
# ## Create new sample of observed values
# newdat <- lapply(newdat, function(x, n = 50) {
#   x[, probFixation := mean(rbinom(n, 1, Fixations)), by = .(Subject, Time)]
# })
# newdat <- rbindlist(newdat)

makenewdat <- function(x, y, nn = 50) {
  newdat <- Map(function(x, y) {
    x[, Fixations := eyetrackSim:::logistic_f(y, Time)]
  }, split(ci, by = "Subject"), split(mm, row(mm)))
  newdat <- lapply(newdat, function(x, n = nn) {
    x[, probFixation := mean(rbinom(n, 1, Fixations)), by = .(Subject, Time)]
  })
  newdat <- rbindlist(newdat)
}

newdat <- makenewdat(ci, mm, 50)


wts <- seq(from = -2, to = 2, length.out = length(TIME))
wts <- dnorm(wts)
earlyT <- sort(sample(TIME, 50, prob = wts))
lateT <- sort(sample(TIME, 50, prob = rev(wts)))

tt <- split(newdat, by = "Subject")
tt[[1]] <- tt[[1]][Time %in% earlyT]
tt[[2]] <- tt[[2]][Time %in% earlyT]
tt[[3]] <- tt[[3]][Time %in% lateT]
tt[[4]] <- tt[[4]][Time %in% lateT]
newdat <- rbindlist(tt)

fit <- bdotsFit(data = newdat, 
                subject = "Subject", 
                time = "Time", 
                y = "probFixation", 
                group = "protocol", 
                curveType = logistic())
#plot(fit)

fv <- lapply(split(fit, by = "Subject"), function(x) {
  data.table(sub = x$Subject, fv = fitted.values(x$fit[[1]]))
})
fv <- rbindlist(fv)
fv[, Time := c(earlyT, earlyT, lateT, lateT)]
fv[, probFixation := newdat$probFixation]
ggplot(fv, aes(x = Time, y = fv, color = sub)) +
  geom_line() + 
  geom_line(aes(x = Time, y = probFixation, color = sub), linetype = "dashed", size = 1) + 
  facet_wrap(~sub)

coef(fit)
mm
```


## {-}
